A Switching Planner for Combined Task and Observation Planning

### 2010-11-24 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
NOt more "blocks". Also, I got rid of the ambiguity with uninitialised variables.

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(:init)$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.g., for
$\pp{DORA}$ $s_0 \equiv \{(= (\pp{is-in}~\pp{DORA})~\pp{kitchen})\}$.

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(:init)$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $s$ effects a transition to a
successor state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with
probability $1 - p_i$. Here, $s^{T_i}$ is the union of $s$ and atomic
terms from $T_i$, and $s^\bot$ is a sink state -- I.e.,
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not uniformly true in the underlying
(resp. abstract) belief-state, we must make two further adjustments to
the problem posed to the sequential planner. First, if
$\action\in\actions$ has a precondition that is not true with
probability $1$ in $\bstate_0$, then before it is executed blindly,
the sequential continual planning switches to a decision-theoretic
session, discussed in detail below. Overall, by scheduling $\action$
the sequential planner makes a naive assumption about the possibility
of the agent eventually ``committing'' to the truth values of specific
state propositions. Thus, the role of decision-theoretic plans is to
examine the validity of assumptions about the runtime state of
sequential planning. Second, if the truth value of $\prop$ is not
known, and no assumption is made about it in the current abstract
state, then if $\prop \in
\poss(\action)$ we have that action $\action$ is not applicable.

@@

In a decision-theoretic session, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ block
from the original problem description.

@@

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ block.  We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the {\em assumptive} actions in the current
sequential plan. For example, if $\pp{DORA}$'s plan is:

@@

\noindent Then the {\em relaxed} block will be,

@@

In a second step, we iteratively refine the {\em relaxed} block by
adding terms from the original block while the resulting state space
remains of a practicable size. Of course, it is expensive to measure
size exactly, and in practice we estimate this to be
$2^{|\propositions^\#|}$, where $\propositions^\#$ is the set of
propositions in the abstraction that were not assigned the {\em
unknown} value \#. Terms are proposed for addition because they are
probabilistic, and because they were omitted in the {\em relaxed}
visitation characterised by {\em assumptive} actions in the sequential
plan. Each iteration, we add one candidate term with the largest
number of parent terms already in the candidate tree.

>>>>

\noindent A valid assumption in our setting corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this relaxed
case, a conjunctive term is visited iff its atomic subterms are
visited, and zero or more of its immediate probabilistic subterms are
visited. The starting state, $\state_0$, for sequential planning is
the unique abstract state with probability $1.0$ -- E.g., for
$\pp{DORA}$:

\[
\begin{array}{l}
s_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $p_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..p_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $s$ effects a transition to a successor
state $s^{T_i}$ with probability $p_i$, and $s^\bot$ with probability
$1 - p_i$. Here, $s^{T_i}$ is the union of $s$ with atomic terms from
$T_i$ and conflicting propositions from $s$ removed. For example, if
the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
delete $(=(\pp{is-in}~\pp{box})~\#)$. State $s^\bot$ is a sink, where
$\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.

Because $\state_0$ is abstract, and because actions can have
preconditions that are not uniformly true in the underlying
(resp. abstract) belief-state, we must make two further adjustments to
the problem posed to the sequential planner. First, if
$\action\in\actions$ has a precondition that is not true with
probability $1$ in $\bstate_0$, then before it is executed blindly,
the sequential continual planning switches to a decision-theoretic
session, discussed in detail below. Overall, by scheduling $\action$
the sequential planner makes a naive assumption about the possibility
of the agent eventually ``committing'' to the truth values of specific
state propositions in $\poss{\action}$. The role of decision-theoretic
plans is to examine the validity of assumptions about the runtime
state of sequential planning. Second, if the truth value of $\prop$ is
not known, and no assumption is made about it in the current abstract
state, then if $\prop \in
\poss(\action)$ we have that action $\action$ is not applicable.

@@

In a decision-theoretic session, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ declaration
from the original problem description.

@@

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration.  We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the {\em assumptive} actions in the current
sequential plan. For example, if $\pp{DORA}$'s plan is:

@@

\noindent Then the {\em relaxed} declaration will be,

@@

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. Each iteration, we add one candidate
term with the largest number of parent terms already in the candidate
tree.
